### 2021-11-07 Final Doc/Chapters/Chapter1.tex
Updated Final Docs

\section{Project Description}
A student's learning experience is typically affected by their enjoyment of interacting with the material and their general interest in the topic material presented. As such, a student will have their own individual preferences and style of learning. Due to the rapid and prevalent development and accessibility of technology, using digital computer games as a means of teaching has been a topic of discussion. In light of this, an academic study will be undertaken as an individual research project to determine how computer games can be used in education. This includes determining what qualities are needed for a digital computer game to be effective as a means of teaching and learning.
\\\\
Additionally, this project aims to create an artefact in the form of a digital computer game as a group practical effort. This artefact will be developed alongside the individual research projects of the group members. It should be noted that this artefact is developed as a standalone project with no links to the research projects of the group members but may be influenced by each members individual research projects as the research may help develop certain aspects of the artefact.
\pagebreak
\section{Project Contents}
\subsection{Problem Description and Background}
Education as it stands is still built on a system that is no longer needed in modern society. Ackoff and Greenberg (2008) explain that the current traditional methods of teaching are no longer as relevant as they once were as it is aimed to produce members of society that were likely to not question any fundamental aspects of how things operated. It is largely a system that focuses on teaching while disregarding learning as the last major stride in development in education was to industrialise it – having them operate efficiently like factories. Ackoff (1991) discusses that while learning is the advancement of one’s understanding and knowledge on a given topic or subject, this can happen in the complete absence of teaching, which is the explanation of a subject from one person to another. The opposite of this is also true, teaching may occur with no learning.
\\\\
One major flaw with this system currently is that it stifles the creativity and drive of some students as each level of education is largely the same and as such monotonous. As such, education needs some form of system to create an interest in learning for the students or risk providing only a means of teaching without individual students learning (\cite{Ackoff2008}). It is therefore vital to provide learners in all levels of education with engaging content or methods of delivery that they will enjoy will cause them to be more motivated to learn and look further into that specific topic (\cite{Ackoff2008}). 
\\\\
With the recent developments in technology and the fact that technology, in general, becoming more accessible, some institutions have adopted some forms of digital learning or assist traditional teaching with digital assistance. Deshpande and Huang (2011) state that the current generation of students is the first to grow up with abundant access to technology. They continue to state that, on average, these students spend almost double the time playing video games as they do reading (\cite{Deshpande2011}). Virvou, Katsionis and Manos (2005) echo the point that computer games are popular among individuals who are in schools and as such could provide a means to deliver content in an interesting and engaging manner. As such, the motivation behind this study is to further investigate the possibility of using video games as a means to encourage learning in teaching environments as current means of teaching may not be optimal for some individuals.

@@

\section{Procedures and Methods of Investigation}
\subsection{Research Paradigm and Methodology Used}
The research aspect of this project will be conducted in accordance with the interpretivist paradigm. The main undertaking of this paradigm is to understand the world and its’ various aspects according to the subjective experiences of people (\cite{kivunja2017understanding}). It is due to this reason that this project uses this paradigm as what each student requires to efficiently learn is subjective to each individual. Kivunja and Kuyini  (2017) further mention that the interpretivist paradigm makes use of subjective epistemology which allows the researcher to analyse the data, which in this project is the principles and qualities that allow for a game to be effective in a learning environment, through their individual cognitive processes. 
\\\\
Kivunja and Kuyini (2017) continue to list certain characteristics of research conducted under this paradigm typically possess. These include the understanding that the social world cannot be fully comprehended from one standpoint which is why this project makes use of various other studies in the compilation of the aforementioned principles and qualities. Another is that knowledge and its’ understanding is developed through the findings of the study itself.
\pagebreak
Design science is a methodology composed of using analytical techniques and can be used according to many paradigms to perform research in the information systems field (\cite{vaishnavi2004design}). This means of research centres on the development of artefacts to better understand certain aspects through the creation of new knowledge and that assessment of an artefacts performance (\cite{vaishnavi2004design}). This methodology was chosen as it provides certain outlines that will allow for the development and analyses of the principles and qualities for a game to be used academically through the creation of such artefact incorporation these qualities into its’ design and development. 
\\\\
Design Science also provides a general structure as to how this project will be completed as shown in Figure \ref{des} below taken from Vaishnavi and Kuechler’s (2004) work. This process allows for the development of an artefact alongside the collection of information and knowledge that can be incorporated into the design and development of the artefact as well as the evaluation of the implementation of the qualities onto the artefact.

@@

A more detailed breakdown of the project plan is presented in Figure 4 below with provisional dates for the deliverables and the anticipated objectives required to meet them.

>>>>

\section{Project Description}
A student's learning experience is typically affected by their enjoyment of interacting with the material and their general interest in the topic material presented. As such, a student will have their own individual preferences and style of learning. Due to the rapid and prevalent development and accessibility of technology, using digital computer games as a means of teaching has been a topic of discussion. In light of this, an academic study will be undertaken as an individual research project to determine how computer games can be used in education. This includes determining what qualities are needed for a digital computer game to be effective as a means of teaching and learning.
\\\\
Additionally, this project aims to create an artefact in the form of a digital computer game as a group practical effort. This artefact will be developed alongside the individual research projects of the group members. It should be noted that this artefact is developed as a standalone project with no links to the research projects of the group members but may be influenced by each members individual research projects as the research may help develop certain aspects of the artefact.
\pagebreak
\section{Project Contents}
\subsection{Problem Description and Background}
Education as it stands is still built on a system that is no longer needed in modern society. Ackoff and Greenberg (2008) explain that the current traditional methods of teaching are no longer as relevant as they once were as it is aimed to produce members of society that were likely to not question any fundamental aspects of how things operated. It is largely a system that focuses on teaching while disregarding learning as the last major stride in development in education was to industrialise it – having them operate efficiently like factories. Ackoff (1991) discusses that while learning is the advancement of one’s understanding and knowledge on a given topic or subject, this can happen in the complete absence of teaching, which is the explanation of a subject from one person to another. The opposite of this is also true, teaching may occur with no learning.
\\\\
One major flaw with this system currently is that it stifles the creativity and drive of some students as each level of education is largely the same and as such monotonous. As such, education needs some form of system to create an interest in learning for the students or risk providing only a means of teaching without individual students learning (\cite{Ackoff2008}). It is therefore vital to provide learners in all levels of education with engaging content or methods of delivery that they will enjoy will cause them to be more motivated to learn and look further into that specific topic (\cite{Ackoff2008}). 
\\\\
With the recent developments in technology and the fact that technology, in general, is becoming more accessible, some institutions have adopted some forms of digital learning or assist traditional teaching with digital assistance. Deshpande and Huang (2011) state that the current generation of students is the first to grow up with abundant access to technology. They continue to state that, on average, these students spend almost double the time playing video games as they do reading (\cite{Deshpande2011}). Virvou, Katsionis and Manos (2005) echo the point that computer games are popular among individuals who are in schools and as such could provide a means to deliver content in an interesting and engaging manner. As such, the motivation behind this study is to further investigate the possibility of using video games as a means to encourage learning in teaching environments as current means of teaching may not be optimal for some individuals.

@@

\section{Procedures and Methods of Investigation}
\subsection{Research Paradigm and Methodology Used}
The research aspect of this project will be conducted in accordance with the interpretivist paradigm. The main undertaking of this paradigm is to understand the world and its’ various aspects according to the subjective experiences of people (\cite{kivunja2017understanding}). It is due to this reason that this project uses this paradigm as what each student requires to efficiently learn is subjective to each individual. Kivunja and Kuyini  (2017) further mention that the interpretivist paradigm makes use of subjective epistemology which allows the researcher to analyse the data, which in this project is the principles and qualities that allow for a game to be effective in a learning environment, through their individual cognitive processes. 
\\\\
Kivunja and Kuyini (2017) continue to list certain characteristics of research conducted under this paradigm typically possess. These include the understanding that the social world cannot be fully comprehended from one standpoint which is why this project makes use of various other studies in the compilation of the aforementioned principles and qualities. Another is that knowledge and its’ understanding is developed through the findings of the study itself.
\newpage
\noindent Design science is a methodology composed of using analytical techniques and can be used according to many paradigms to perform research in the information systems field (\cite{vaishnavi2004design}). This means of research centres on the development of artefacts to better understand certain aspects through the creation of new knowledge and that assessment of an artefacts performance (\cite{vaishnavi2004design}). This methodology was chosen as it provides certain outlines that will allow for the development and analyses of the principles and qualities for a game to be used academically through the creation of such artefact incorporation these qualities into its’ design and development. 
\\\\
Design Science also provides a general structure as to how this project will be completed as shown in Figure \ref{des} below taken from Vaishnavi and Kuechler’s (2004) work. This process allows for the development of an artefact alongside the collection of information and knowledge that can be incorporated into the design and development of the artefact as well as the evaluation of the implementation of the qualities onto the artefact. In  this case, the artefact will take the form of an article.

@@

\noindent A more detailed breakdown of the project plan is presented in Figure 4 below with provisional dates for the deliverables and the anticipated objectives required to meet them.

### 2021-11-07 Final Doc/Chapters/Chapter2.tex
Updated Final Docs

This current paradigm of education is operated on the basis of standardisation to accommodate the change in societal needs and requirements that the industrial age brought with it (\cite{Reigeluth1996}). Reigeluth (1996) states that this system was developed on the fundamental need of sorting students into different categories and to allow for comparisons so that the training of the workforce could also be separated into different groups, those being labourers, having “hands-on” skills, and managers, who mainly had a managerial skill set. 
\\\\
However, with the shift into the “information age”, the requirements of the workforce have also changed. The industrial age being a time of mass-production with the emergence of various new processing technologies and the information age being characterised by the fact that information is being transmitted and generated at an ever-increasing rate due to further technological developments (\cite{gibson2006games, Reigeluth1996}). The most notable changes from the aforementioned paradigm shifts are that from conformity and compliance to initiative and diversity where greater value is placed on each individual’s strength and contribution to a project or organisation rather than each individual having the same skill set (\cite{Reigeluth1996}).  A few other differences between the needs of organisations during the industrial age and the new requirements brought with the information age are depicted in the above table.
\newpage
\noindent Reigeluth (1996) continues and states that the current paradigm is not focused on learning but rather categorisation. Ackoff (1991) hold a similar viewpoint stating that there is more of a focus on teaching rather than learning. It should be noted that teaching and learning are very distinct from one another as both can take place without the other (\cite{Ackoff1991}). Learning is defined as increasing one’s ability to perform an act effectively while trying to meet an objective through acquiring new knowledge whereas teaching is the process of providing this knowledge (\cite{Ackoff1991}). Through this, it is clear that institutions under this paradigm aim to give learners a verbose vocabulary to speak on topics that they do not fully comprehend (\cite{Ackoff1991}). 
\\\\
Due to this aforementioned paradigm shift and in what requirements are desired by most organisations in the information age, a shift in instructional theory is also needed – one such as going from making use of passive learning through traditional teaching means to one that is centred on active learning (\cite{Reigeluth1996}). These changes are discussed in more detail in the following section on pedagogy and learning theories and the means to accomplish them in the further sections on ludology and gamification.

>>>>

\noindent This current paradigm of education is operated on the basis of standardisation to accommodate the change in societal needs and requirements that the industrial age brought with it (\cite{Reigeluth1996}). Reigeluth (1996) states that this system was developed on the fundamental need of sorting students into different categories and to allow for comparisons so that the training of the workforce could also be separated into different groups, those being labourers, having “hands-on” skills, and managers, who mainly had a managerial skill set. 
\\\\
However, with the shift into the “information age”, the requirements of the workforce have also changed. The industrial age being a time of mass-production with the emergence of various new processing technologies and the information age being characterised by the fact that information is being transmitted and generated at an ever-increasing rate due to further technological developments (\cite{gibson2006games, Reigeluth1996}). The most notable changes from the aforementioned paradigm shifts are that from conformity and compliance to initiative and diversity where greater value is placed on each individual’s strength and contribution to a project or organisation rather than each individual having the same skill set (\cite{Reigeluth1996}).  A few other differences between the needs of organisations during the industrial age and the new requirements brought with the information age are depicted in the above table.
\newpage
\noindent Reigeluth (1996) continues and states that the current paradigm is not focused on learning but rather categorisation. Ackoff (1991) hold a similar viewpoint stating that there is more of a focus on teaching rather than learning. It should be noted that teaching and learning are very distinct from one another as both can take place without the other (\cite{Ackoff1991}). Learning is defined as increasing one’s ability to perform an act effectively while trying to meet an objective through acquiring new knowledge whereas teaching is the process of providing this knowledge (\cite{Ackoff1991}). Through this, it is clear that institutions under this paradigm aim to give learners a verbose vocabulary to speak on topics that they do not fully comprehend (\cite{Ackoff1991}). 
\\\\
Due to this aforementioned paradigm shift and in what requirements are desired by most organisations in the information age, a shift in instructional theory is also needed – one such as going from making use of passive learning through traditional teaching means to one that is centred on active learning (\cite{Reigeluth1996}). These changes are discussed in more detail in the following section on pedagogy and learning theories and the means to accomplish them in the further sections on ludology and gamification.

### 2021-11-07 Final Doc/Chapters/Chapter3.tex
Updated Final Docs

\noindent For the first aspect of this project, the resulting outcome was an academic article titled, \textit{"Linking Gamification, Ludology and Pedagogy: How to Use Serious Games for Various Knowledge Domains"}. As stipulated in the title, the fields of gamification, ludology and pedagogy were the focus of this research aspect of the project. In this article, the question posed by this project is again reiterated and sequentially answered. The article is attached to this document under Appendix  \ref{AppendixC}. 
\\\\
The artefact accompanying this research project is a digital computer came which was developed as part of a group effort as mentioned previously. While this artefact is separate from the individual research studies conducted by its' group members, it may have been influenced by the topics being researched. The artefact is titled \textit{"Puzzle Ball: Spherical Shadows"} and is further categorised as a 3D puzzle platformer with aspects of action gameplay played in first person. A playable version of the artefact is available on the website itch.io website\footnote{\url{https://josh-scg.itch.io/puzzle-ball-spherical-shadows}} and the source code of the artefact is available on GitHub\footnote{\url{https://github.com/GCWehmeyer/Spherical_Shadows}}$^{,}$\footnote{\url{https://github.com/Josh-SCG/Spherical_Shadows}}

@@

\noindent

\section{Description of the Development of the Artefact}
%Comprehensive report on the way in which each phase of the life cycle was applied in the development of this artefact.

\subsection{Sprint 1: Know the Basics}

\begin{figure}[H]
\centering
\includegraphics[scale=0.33]{Figures/level.png}
\caption{Individual First Level}
\end{figure}

@@

\begin{minted}[
frame=lines,
framesep=2mm,
baselinestretch=1.2,
bgcolor=gray!10,
fontsize=\footnotesize,
breaklines, 
linenos,
]{csharp}
/*OnFloor() checks if player is on ground
     * part of ground detection system
     */
    private bool OnFloor() 
    {
        //array of raycasts? to check if touching "floor"
        Collider[] hitColliders = Physics.OverlapBox(transform.position, transform.localScale / 3f, Quaternion.Euler(0, 0, 0), floorMask);
        bool check = false;
        //cycle through array >> if none true then not on floor, if just one, on floor
        for (int i = 0; i < hitColliders.Length; i++)
        {
            if (hitColliders[i]) 
            {
                check = true;
            }
        }
        return check;
    }
\end{minted}

@@

>>>>

\noindent For the first aspect of this project, the resulting outcome was an academic article titled, \textit{"Linking Gamification, Ludology and Pedagogy: How to Use Serious Games for Various Knowledge Domains"}. As stipulated in the title, the fields of gamification, ludology and pedagogy were the focus of this research aspect of the project. In this article, the question posed by this project is again reiterated and sequentially answered. The article is attached to this document under Appendix  \ref{AppendixC}. 
\\\\
The artefact accompanying this research project is a digital computer came which was developed as part of a group effort as mentioned previously. While this artefact is separate from the individual research studies conducted by its' group members, it may have been influenced by the topics being researched. The artefact is titled \textit{"Puzzle Ball: Spherical Shadows"} and is further categorised as a 3D puzzle platformer with aspects of action gameplay played in first person. A playable version of the artefact is available on the website itch.io\footnote{\url{https://josh-scg.itch.io/puzzle-ball-spherical-shadows}} and the source code of the artefact is available on GitHub\footnote{\url{https://github.com/GCWehmeyer/Spherical_Shadows}}$^{,}$\footnote{\url{https://github.com/Josh-SCG/Spherical_Shadows}}

@@

\newpage

\section{Description of the Development of the Artefact}
This section will detail the development of the computer game artefact but will only include specifics on the aspects which were worked on individually - as such the contributions and accompanying processes of the other group members will not be discussed.

\subsection{Sprint 1: Know the Basics}
The initial sprint for development of the artefact had a main focus on each group member learning the basics of both game design in general as well the how to use the Unity Editor, specifically the 2019.4.22f version. The reason for going with this build of the editor was due to the fact that at the time, it was one that was targeted for long term support by Unity. 
\\\\
The aim of learning the basics of game design saw each member in the group informally gathering information on various related topics within game design. This included topics such as level design and the psychology used during game design in addition to examining so-called "Developer Logs", which is typically a video presentation on the progress made at certain increments during the overall development of a game, by other game designers to gain insight into how the structure of game development typically flows. 
\\\\
As a part this process, each member of the group was tasked with developing a singular level. This was done with the aim to gain practical experience with the Unity Editor as well as serve as a means for each group member to discuss what had been discovered while gathering insight into game design and share that acquired knowledge with the other members. Figure \ref{1lvl} shows one of the levels created.

\begin{figure}[H]
\centering
\includegraphics[scale=0.33]{Figures/level.png}
\caption{Individual First Level}
\label{1lvl}
\end{figure}

\noindent The resources made use of for this particular sprint, those being any material used in the knowledge gathering phase, were mostly found through a publicly available online excel sheet\footnote{\url{https://docs.google.com/spreadsheets/d/1QhFyPfYSjHv7PjibGrslF3mNW_CIDXWv9o-iQgLbu1o}} which contains information of various aspects of game design.
\\\\
This sprint was concluded upon the discussion of what each group member had learnt and with the next sprint being planned. 
\\\\
The key aspects from this sprint's knowledge gathering phase were that the ProBuilder tool kit within the Unity Editor would be used for the construction and development of levels moving forward as it allowed for the creation on a singular large object which could be adjusted and shaped. Using this one object in place of many smaller ones arranged in the same shape allowed for a less computational intensive process when the application is run due to the way a scene is loaded. 
\\\\
Additionally, the character controller, that being how the in-game character is affected by a user's inputs, would be developed from the ground up in place of using the one pre-built by Unity as it allows for the most customisation of features.

@@

@@

\begin{figure}[H]
\centering
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[scale=1.4]{Figures/crossold.png}
  \caption{Place holder Crosshair}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
  \centering
  \includegraphics[scale=1]{Figures/crosshair.png}
  \caption{Finalised Crosshair made with GIMP}
\end{subfigure}
\caption{Crosshair Development}
\end{figure}

### 2021-11-07 Final Doc/Chapters/Chapter5.tex
Updated Final Docs

\section{What was Learnt Throughout this Project}
In terms of what was learnt during the duration of this project, there were several new technologies and software in addition to a vast amount of knowledge from various fields.
\\\\
On the software side of things, the largest one was the Unity Editor, specifically the 2019.4.22f build of it, which was the main tool in developing the artefact. The choice of using Unity over another game engine, such as Unreal or Godot, was mainly due to the programming language it used (that being C\#). This was the language that most of the group, excluding myself, was familiar with and most experienced with when compared to the alternatives. Following this, C\# was also a new language learnt for this project and while the initial usage of it was difficult, with the experience  Java, another object-oriented language, it became easy to write and implement the scripts needed. 
\\\\
Another technology made use of, and consequently learnt, is GitHub. This was used for version control and was the simplest way to handle it. The GitHub application was used for facilitating this. 
\\\\
\LaTeX was another technology used in the latter half the project. It is a software tool that allows for the creation of PDF's and was used for the Article as well as the final documentation. While it was initially hard to pick up and use, it also became easier with time. It allowed for these respective documents to be presented very neatly and should have been used earlier on in the project's development. 
\\\\
A fair amount of information concerning game design was also learnt form how psychology plays a factor in the choices game designers take when developing a game or user interface. While a fair amount of this knowledge was used when designing the artefact, there is still much more that could have been incorporated - such as the aforementioned use of psychology to style the game better.
\\\\
Additionally, various knowledge in the fields used for the research article was learnt. These were fields concerning educational and instructional theories, the study of games as well as turning tasks into some form of game be it digital or not. These topics were interesting to examine and read about and were a tremendous help on how to approach any future educational endeavours such as tutoring. 
\\\\
As this project was described to be under \textit{Entertainment Computing}, being able to analyse other computer games was vital to garner a proper understanding of how to go about the development of both the artefact and article. This process was heavily aided by Prof. Günther Drevin where we were tasked with various presentations on the field of ludology itself and on the analysis of a game - Myst.

>>>>

\section{What was Learnt Throughout this Project}
In terms of what was learnt during the duration of this project, there were several new technologies and software in addition to a vast amount of knowledge from various fields.
\\\\
On the software side of things, the largest one was the Unity Editor, specifically the 2019.4.22f build of it, which was the main tool in developing the artefact. The choice of using Unity over another game engine, such as Unreal or Godot, was mainly due to the programming language it used (that being C\#). This was the language that most of the group, excluding myself, was familiar with and most experienced with when compared to the alternatives. Following this, C\# was also a new language learnt for this project and while the initial usage of it was difficult, with the experience  Java, another object-oriented language, it became easy to write and implement the scripts needed. 
\\\\
Another technology made use of, and consequently learnt, is GitHub. This was used for version control and was the simplest way to handle it. The GitHub Desktop application was used for facilitating this. 
\\\\
\LaTeX was another technology used in the latter half the project. It is a software tool that allows for the creation of PDF's and was used for the Article as well as the final documentation. While it was initially hard to pick up and use, it also became easier with time. It allowed for these respective documents to be presented very neatly and should have been used earlier on in the project's development. 
\\
A fair amount of information concerning game design was also learnt form how psychology plays a factor in the choices game designers take when developing a game or user interface. While a fair amount of this knowledge was used when designing the artefact, there is still much more that could have been incorporated - such as the aforementioned use of psychology to style the game better.
\\\\
Additionally, various knowledge in the fields used for the research article was learnt. These were fields concerning educational and instructional theories, the study of games as well as turning tasks into some form of game be it digital or not. These topics were interesting to examine and read about and were a tremendous help on how to approach any future educational endeavours such as tutoring. 
\\\\
As this project was described to be under \textit{Entertainment Computing}, being able to analyse other computer games was vital to garner a proper understanding of how to go about the development of both the artefact and article. This process was heavily aided by Prof. Günther Drevin where we were tasked with various presentations on the field of ludology itself and on the analysis of a game - Myst.
